TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2018-08-21 paper/draft.tex
fix alignment
non-ascii character

\subsection{Temperature/Metallicity Model \label{subsec:mann_results}}
For the physical parameter model we trained \thecannon\ on 87 M dwarfs with a two-dimensional temperature/metallicity labels, to a precision of 77K/0.09dex as estimated by the cross validation scatter, similar to the original training sample uncertainties of 60K/0.08dex. We note for this model 6 out of 87 sources show strong visible rotational line broadening (as indicated by the red circles in Figure \ref{fig:mann_validation}), while the remaining sources are mostly old ($>8$ Gyr) and likely slowly rotating sources. While these sources with high rotational broadening have high $\chi^2$ values (those sources with $\chi^2 > 80,000$ in Figure \ref{fig:chi_dist}), we find that the scatter of labels derived for these sources is consistent with the rest of the sample and we do not remove them from the training sample.

@@

\begin{figure}[ht]
\plotone{figures/self_and_validation_test_teff_snr.png}
\plotone{figures/self_and_validation_test_feh_snr.png}
\caption{Consistency test for the Mann-trained temperature/metallicity type model; self-test (left), and leave-one-out cross validation (right). Sources with strong rotational broadening (2M07171706−0501031, 2M08155393+3136392, 2M09174473+4612246, 2M10331367+3409120, 2M15594729+4403595, 2M13400879+4346380), as identified by visual inspection are marked with red circles. Predictive accuracy, as computed by the scatter in cross validation is 77K in temperature, and 0.09 dex in metallicity.} \label{fig:mann_validation}
\end{figure}

\begin{figure}[ht]
\begin{center}
\plotone{figures/self_and_validation_test_spt_snr.png}
\end{center}
\caption{Consistency test for the West-trained spectral type model; self-test (left), and leave-one-out cross validation (right).  Predictive accuracy, as computed by the scatter in cross validation is 0.9 subtypes.} \label{fig:west_validation}
\end{figure}

%---------------------------
\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/demo_derivatives_teff1.png}
\includegraphics[width=16cm]{figures/demo_derivatives_teff2.png}
\includegraphics[width=16cm]{figures/demo_derivatives_teff3.png}
\end{center}
\caption{\textit{Top two panels of each plot:} Mann-trained model for varying temperatures, and similar metallcities. \textit{Third panel of each plot:} Derivative of \thecannon\ model with respect to temperature, taken at the median training temperature, T$_{\rm eff}=3463$K; error estimate computed using a jackknife statistic at each pixel is marked in red, making it possible to distinguish which features vary significantly with change in spectral type, and which are likely due to noise.} \label{fig:demo_teff}
\end{figure}

@@

\begin{deluxetable*}{rcl}
\tablecaption{Test results table header description.}
\tablehead{\colhead{Column} & \colhead{Unit} & \colhead{Description}}
\startdata
{\tt APOGEE$\_$ID} 	& 	& \apogee\ \zmass\ designation \\
{\tt BP$\_$MAG }		& mag	& \gaia\ $BP$ band photometry	\\
{\tt BP$\_$RP} 		& mag	& \gaia\ $BP-RP$ color	\\
{\tt CHI$\_$MANN} 	& 	& $\chi^2$ fit of Mann-trained \cannon\ model	\\
{\tt CHI$\_$WEST} 	& 	& $\chi^2$ fit of West-trained \cannon\ model	\\
{\tt DEC} 			& deg	& \apogee\ declination angle	\\
{\tt DIST} 			& pc	& Distance (1/$\varpi$)	\\
{\tt FE$\_$H }		& dex	& Mann-trained \cannon\ [Fe/H]	\\
{\tt GAIA$\_$ID} 		&	& \gaia\ identification number \\
{\tt G$\_$ABS} 		& mag	& \gaia\ $G$ band absolute magnitude	\\
{\tt G$\_$MAG} 		& mag	& \gaia\ $G$ band photometry	\\
{\tt H} 				& mag	& \apogee\ $H$ band photometry	\\
{\tt H$\_$ERR} 		& mag	& \apogee\ $H$ band photometry uncertainty	\\
{\tt J} 				& mag	& \apogee\ $J$ band photometry	\\
{\tt J$\_$ERR} 		& mag	& \apogee\ $J$ band photometry uncertainty	\\
{\tt K} 				& mag	& \apogee\ $K$ band photometry	\\
{\tt K$\_$ERR} 		& mag	& \apogee\ $K$ band photometry uncertainty	\\
{\tt PLX} 			& mas	& \gaia\ parallax	\\
{\tt PLX$\_$ERR} 		& mas	& \gaia\ parallax uncertainty	\\
{\tt PMDEC} 			& mas yr$^{-1}$	& \gaia\ declination proper motion	\\
{\tt PMDEC$\_$ERR} 	& mas yr$^{-1}$	& \gaia\ declination proper motion uncertainty	\\
{\tt PMRA} 			& mas yr$^{-1}$	& \gaia\ radial ascension proper motion	\\
{\tt PMRA$\_$ERR} 	& mas yr$^{-1}$	& \gaia\ radial ascension proper motion uncertainty	\\
{\tt RA} 				& deg	& \apogee\ radial ascension angle	\\
{\tt RP$\_$MAG} 		& mag	& \gaia\ $RP$ band photometry	\\
{\tt RV$\_$APOGEE} 	& km s$^{-1}$	& \apogee\ radial velocity	\\
{\tt RV$\_$APOGEE$\_$ERR} 	& km s$^{-1}$	& \apogee\ radial velocity uncertainty	\\
{\tt RV$\_$GAIA} 				& km s$^{-1}$	& \gaia\ radial velocity	\\
{\tt RV$\_$GAIA$\_$ERR}		& km s$^{-1}$	& \gaia\ radial velocity uncertainty	\\
{\tt SPT} 				& 	& West-trained \cannon\ spectral subtype ($M0-M9$)	\\
{\tt TEFF} 				& K	& Mann-trained \cannon\ effective temperature	\\
{\tt Vx} 				& km s$^{-1}$	& Cartesian $x$ velocity in Galactocentric coordnates	\\
{\tt Vy} 				& km s$^{-1}$	& Cartesian $y$ velocity in Galactocentric coordnates	\\
{\tt Vz} 				& km s$^{-1}$	& Cartesian $z$ velocity in Galactocentric coordnates	\\
{\tt X} 				& pc	& Cartesian $x$ position in Galactocentric coordnates	\\
{\tt Y} 				& pc	& Cartesian $y$ position in Galactocentric coordnates	\\
{\tt Z} 				& pc	& Cartesian $z$ position in Galactocentric coordnates	
\enddata
\end{deluxetable*}

>>>>

\subsection{Temperature/Metallicity Model \label{subsec:mann_results}}
For the physical parameter model we trained \thecannon\ on 87 M dwarfs with a two-dimensional temperature/metallicity labels, to a precision of 77K/0.09dex as estimated by the cross validation scatter, similar to the original training sample uncertainties of 60K/0.08dex. We note for this model 6 out of 87 sources show strong visible rotational line broadening (as indicated by the red circles in Figure~\ref{fig:mann_validation}), while the remaining sources are mostly old ($>8$ Gyr) and likely slowly rotating sources. While these sources with high rotational broadening have high $\chi^2$ values (those sources with $\chi^2 > 80,000$ in Figure \ref{fig:chi_dist}), we find that the scatter of labels derived for these sources is consistent with the rest of the sample and we do not remove them from the training sample.

@@

\begin{figure}[ht]
\plotone{figures/self_and_validation_test_teff_snr.png}
\plotone{figures/self_and_validation_test_feh_snr.png}
\caption{Consistency test for the Mann-trained temperature/metallicity type model; self-test (left), and leave-one-out cross validation (right). Sources with strong rotational broadening (2M07171706--0501031, 2M08155393+3136392, 2M09174473+4612246, 2M10331367+3409120, 2M15594729+4403595, 2M13400879+4346380), as identified by visual inspection are marked with red circles. Predictive accuracy, as computed by the scatter in cross validation is 77K in temperature, and 0.09 dex in metallicity.}\label{fig:mann_validation}
\end{figure}

\begin{figure}[ht]
\begin{center}
\plotone{figures/self_and_validation_test_spt_snr.png}
\end{center}
\caption{Consistency test for the West-trained spectral type model; self-test (left), and leave-one-out cross validation (right).  Predictive accuracy, as computed by the scatter in cross validation is 0.9 subtypes.}\label{fig:west_validation}
\end{figure}

%---------------------------
\begin{figure}[ht]
\begin{center}
\includegraphics[width=16cm]{figures/demo_derivatives_teff1.png}
\includegraphics[width=16cm]{figures/demo_derivatives_teff2.png}
\includegraphics[width=16cm]{figures/demo_derivatives_teff3.png}
\end{center}
\caption{\textit{Top two panels of each plot:} Mann-trained model for varying temperatures, and similar metallcities. \textit{Third panel of each plot:} Derivative of \thecannon\ model with respect to temperature, taken at the median training temperature, T$_{\rm eff}=3463$K; error estimate computed using a jackknife statistic at each pixel is marked in red, making it possible to distinguish which features vary significantly with change in spectral type, and which are likely due to noise.}\label{fig:demo_teff}
\end{figure}

@@

\begin{deluxetable*}{rcl}
\tablecaption{Test results table header description.}
\tablehead{Column & Unit & Description}
\startdata
{\tt APOGEE$\_$ID} 	& 	& \apogee\ \zmass\ designation \\
{\tt BP$\_$MAG }		& mag	& \gaia\ $BP$ band photometry	\\
{\tt BP$\_$RP} 		& mag	& \gaia\ $BP-RP$ color	\\
{\tt CHI$\_$MANN} 	& 	& $\chi^2$ fit of Mann-trained \cannon\ model	\\
{\tt CHI$\_$WEST} 	& 	& $\chi^2$ fit of West-trained \cannon\ model	\\
{\tt DEC} 			& deg	& \apogee\ declination angle	\\
{\tt DIST} 			& pc	& Distance (1/$\varpi$)	\\
{\tt FE$\_$H }		& dex	& Mann-trained \cannon\ [Fe/H]	\\
{\tt GAIA$\_$ID} 		&	& \gaia\ identification number \\
{\tt G$\_$ABS} 		& mag	& \gaia\ $G$ band absolute magnitude	\\
{\tt G$\_$MAG} 		& mag	& \gaia\ $G$ band photometry	\\
{\tt H} 				& mag	& \apogee\ $H$ band photometry	\\
{\tt H$\_$ERR} 		& mag	& \apogee\ $H$ band photometry uncertainty	\\
{\tt J} 				& mag	& \apogee\ $J$ band photometry	\\
{\tt J$\_$ERR} 		& mag	& \apogee\ $J$ band photometry uncertainty	\\
{\tt K} 				& mag	& \apogee\ $K$ band photometry	\\
{\tt K$\_$ERR} 		& mag	& \apogee\ $K$ band photometry uncertainty	\\
{\tt PLX} 			& mas	& \gaia\ parallax	\\
{\tt PLX$\_$ERR} 		& mas	& \gaia\ parallax uncertainty	\\
{\tt PMDEC} 			& mas yr$^{-1}$	& \gaia\ declination proper motion	\\
{\tt PMDEC$\_$ERR} 	& mas yr$^{-1}$	& \gaia\ declination proper motion uncertainty	\\
{\tt PMRA} 			& mas yr$^{-1}$	& \gaia\ radial ascension proper motion	\\
{\tt PMRA$\_$ERR} 	& mas yr$^{-1}$	& \gaia\ radial ascension proper motion uncertainty	\\
{\tt RA} 				& deg	& \apogee\ radial ascension angle	\\
{\tt RP$\_$MAG} 		& mag	& \gaia\ $RP$ band photometry	\\
{\tt RV$\_$APOGEE} 	& km s$^{-1}$	& \apogee\ radial velocity	\\
{\tt RV$\_$APOGEE$\_$ERR} 	& km s$^{-1}$	& \apogee\ radial velocity uncertainty	\\
{\tt RV$\_$GAIA} 				& km s$^{-1}$	& \gaia\ radial velocity	\\
{\tt RV$\_$GAIA$\_$ERR}		& km s$^{-1}$	& \gaia\ radial velocity uncertainty	\\
{\tt SPT} 				& 	& West-trained \cannon\ spectral subtype ($M0-M9$)	\\
{\tt TEFF} 				& K	& Mann-trained \cannon\ effective temperature	\\
{\tt Vx} 				& km s$^{-1}$	& Cartesian $x$ velocity in Galactocentric coordnates	\\
{\tt Vy} 				& km s$^{-1}$	& Cartesian $y$ velocity in Galactocentric coordnates	\\
{\tt Vz} 				& km s$^{-1}$	& Cartesian $z$ velocity in Galactocentric coordnates	\\
{\tt X} 				& pc	& Cartesian $x$ position in Galactocentric coordnates	\\
{\tt Y} 				& pc	& Cartesian $y$ position in Galactocentric coordnates	\\
{\tt Z} 				& pc	& Cartesian $z$ position in Galactocentric coordnates	
\enddata
\end{deluxetable*}
